---
title: "PCA with Penguins"
author: "Lizel Potgieter"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load("tidyverse", "palmerpenguins", "corrr", "GGally", "recipes", "tidytext", "tidyr", "here")
here::i_am("penguins_pca.Rmd")

```

This tutorial was taken from https://allisonhorst.github.io/palmerpenguins/articles/pca.html

### Summary reminder
First, we have a little summary as a reminder of what we have in the penguins dataframe
```{r}
summary(penguins)
```
### Correlation Matrix
```{r}
penguins_corr <- penguins %>%
  dplyr::select(body_mass_g, ends_with("_mm")) %>%
  correlate() %>%
  rearrange()
penguins_corr
```

### Pairwise plot matrix
Here we compare the various phenotypes measured from the different penguin species
```{r}
penguins %>%
  select(species, body_mass_g, ends_with("_mm")) %>% 
  GGally::ggpairs(aes(color = species),
          columns = c("flipper_length_mm", "body_mass_g", 
                      "bill_length_mm", "bill_depth_mm")) +
  scale_colour_manual(values = c("darkorange","purple","cyan4")) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"))
```


### PCA
There are many, many different tools that you can use to perform a PCA analysis. See Adam's presentation for even more ways to do this!
```{r}
penguin_recipe <-
  recipe(~., data = penguins) %>% 
  update_role(species, island, sex, year, new_role = "id") %>% 
  step_naomit(all_predictors()) %>% 
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors(), id = "pca") %>% 
  prep()

penguin_pca <- 
  penguin_recipe %>% 
  tidy(id = "pca") 

penguin_pca
```

The value column output here is the loading. This format is not in the right tidy format, so let's fix that!
```{r}
penguin_recipe <-
  recipe(~., data = penguins) %>% 
  update_role(species, island, sex, year, new_role = "id") %>% 
  step_naomit(all_predictors()) %>% 
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors(), id = "pca") %>% 
  prep()

penguin_pca <- 
  penguin_recipe %>% 
  tidy(id = "pca") 

penguin_pca
```

Now we want to see how much each component accounts for in terms of variance

```{r}
penguin_recipe %>% 
  tidy(id = "pca", type = "variance") %>% 
  dplyr::filter(terms == "percent variance") %>% 
  ggplot(aes(x = component, y = value)) + 
  geom_col(fill = "#b6dfe2") + 
  xlim(c(0, 5)) + 
  ylab("% of total variance")
```

And now to plot our penguins PCA!
We need to fix he format.
```{r}
pca_wider <- penguin_pca %>% 
  tidyr::pivot_wider(names_from = component, id_cols = terms)
```

And then prepare our ggplot script
```{r}
# define arrow style
arrow_style <- arrow(length = unit(.05, "inches"),
                     type = "closed")


pca_plot <-
  juice(penguin_recipe) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = species, shape = species), 
             alpha = 0.8, 
             size = 2) +
  scale_colour_manual(values = c("darkorange","purple","cyan4")) 

pca_plot +
  geom_segment(data = pca_wider,
               aes(xend = PC1, yend = PC2), 
               x = 0, 
               y = 0, 
               arrow = arrow_style) + 
  geom_text(data = pca_wider,
            aes(x = PC1, y = PC2, label = terms), 
            hjust = 0, 
            vjust = 1,
            size = 5, 
            color = '#0A537D') 
```

Let's also see PC2 and PC3
```{r}
pca_plot %+% 
  aes(PC2, PC3) +
  geom_segment(data = pca_wider,
               aes(xend = PC2, yend = PC3), 
               x = 0, 
               y = 0, 
               arrow = arrow_style) + 
  geom_text(data = pca_wider,
            aes(x = PC2, y = PC3, label = terms), 
            hjust = 0, 
            vjust = 1,
            size = 5, 
            color = '#0A537D') 
```

